---
title: "How to Run R Codes"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
```

## Two ways to run R codes

Console

R Script

---

## Console

Type the following codes in Console pane of RStudio, line by line.

```{r, eval=FALSE}
mean(mtcars$mpg)
```

```{r, eval=FALSE}
plot(mtcars$hp, mtcars$mpg)
```

Note that R shows the results in the same Console or in the Plots pane.

---

## R Script

Console: difficult to reproduce or revise your analysis later. 

R Script allows you to save the whole codes as a file.

---

Click 'File - New File - R Script`.

In the new Script editor, type the following codes. 

- Press 'ctrl-enter' or 'cmd-return' instead of enter key, *at the end of each line*:

```{r, eval=FALSE}
mean(mtcars$mpg)
```

```{r, eval=FALSE}
plot(mtcars$hp, mtcars$mpg)
```

Save the program as a file in 'Temp' folder, by clicking 'File - Save'.

---

If you want to run multiple lines of codes, select the lines and press 'ctrl-enter' or 'cmd-return'.

--

You can run the whole program by clicking 'Source' button.

- Click 'Session' - 'Restart R' to clear any objects created in previous sessions. 

- Go to the Script editor, and click 'Source' button.


---

Use both Console and R Script for an analysis.

- Include the codes you want to keep, in your R Script.

- Run the codes you do not want to keep, in Console.

---

You are installing tidyverse package using the following command: 

install.packages("tidyverse")

Which one would you use?

A. Console

B. R Script

---

You are running a series of commands for your research projects.

Which one would you use?

A. Console

B. R Script

---

What is the shortcut key when you run each line in R Script?

