---
title: "ggplot2 - 1"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
options(tibble.print_max = Inf)
library(kableExtra)
```

ggplot2 grammar

geom_point()

geom_smooth()

---

```{r eval=FALSE}
install.packages("gapminder")
```
Need to install gapminder package, just once.

```{r}
library(gapminder)
library(tidyverse)
```

No need to load ggplot2 package in addition to tidyverse.

```{r}
gapminder_2007 <- gapminder %>% filter(year == 2007)
```

```{r eval=FALSE}
gapminder_2007
```

---

.scroll-output[
```{r echo=FALSE}
gapminder_2007 %>% kable()
```
]

---

#### Minimum ggplot2 template

```{r eval=FALSE}
ggplot() + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>), data = <DATA>)
```


---

.small[
```{r}
ggplot() + 
  geom_point(mapping=aes(x=gdpPercap, y=lifeExp), data=gapminder_2007)
```
]

---

#### Different geoms

Click Help - Cheatsheets - Data visualization with ggplot2.

Find geom_point().

Which section is it in?

--

Find geom_smooth()

---

.small[
```{r}
ggplot() + 
  geom_smooth(mapping=aes(x=gdpPercap, y=lifeExp), data=gapminder_2007)
```
]

---

#### Overlaying multiple geoms

.small[
```{r}
ggplot() + 
  geom_smooth(mapping=aes(x=gdpPercap, y=lifeExp), data=gapminder_2007) +
  geom_point(mapping=aes(x=gdpPercap, y=lifeExp), data=gapminder_2007)
```
]

---

#### Global option

.small[
```{r eval=FALSE}
ggplot() + 
  geom_smooth(mapping=aes(x=gdpPercap, y=lifeExp), data=gapminder_2007) +
  geom_point(mapping=aes(x=gdpPercap, y=lifeExp), data=gapminder_2007) 
```
]

--

```{r eval=TRUE, fig.height=4.5}
ggplot(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp)) + 
  geom_smooth() +  geom_point()
```

---

#### Matching arguments by position

[? `ggplot`](https://ggplot2.tidyverse.org/reference/ggplot.html)

[? `aes`](https://ggplot2.tidyverse.org/reference/aes.html)

```{r eval=FALSE}
ggplot(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp)) + 
  geom_smooth() +  geom_point()
```

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp ) ) + 
  geom_smooth() + geom_point()
```

---

[`? geom_point`](https://ggplot2.tidyverse.org/reference/geom_point.html)

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp))
```

---

#### Aesthetic mappings

`mapping = aes( <MAPPING> )` maps variables (i.e., columns) in the data to visual properties.

---

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, color=continent))
```

---

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, shape=continent))
```

---

.small[
```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, shape=continent, color=continent))
```
]

---

#### ggplot2 cheatsheet

In RStudio, click Help - Cheatsheets - Data visualization with ggplot2.

--

Note different geom functions may offer different aesthetic properties.

--

Run the example for geom_col()

What are argument names for mpg, aes(class, hwy), class, hwy?

---

```{r include=params$answers}
f <- ggplot(mpg, aes(class, hwy)) 
f + geom_col()
```

---

```{r include=params$answers}
ggplot(mpg, aes(class, hwy)) + geom_col()
```


---

#### Application - Housing value vs. population size across CMAs

Using the PUMF, calculate population and average housing value for each CMA.

Generate a scatter plot with population on x-axis and average housing value on y-axis. (See next slide for the output.)

---

```{r echo=FALSE}
pumf <- read_rds("../Data/2016_census_pumf_ind.rds")
pumf <- pumf %>% mutate(VALUE2 = case_when( VALUE > 80000000 ~ NA, .default = VALUE))

temp <- pumf %>% group_by(CMA) %>% 
  summarise(
    avg_VALUE = weighted.mean(VALUE2, w=WEIGHT, na.rm = TRUE), 
    pop=sum(WEIGHT)
    )

temp <- temp %>% 
  filter(CMA != "Other census metropolitan areas, census agglomerations and other geographies")

temp <- temp %>% arrange( desc(avg_VALUE) )

ggplot() + geom_point(aes(pop, avg_VALUE), temp)
```


---


.small[
Load the pumf dataset.

Add a new column VALUE2 that has NA for VALUE > 80000000 and is equal to VALUE otherwise.(Read chapter 2 VALUE page in the manual to understand why.)

```{r eval=FALSE}
mutate(VALUE2 = case_when( VALUE > 80000000 ~ NA, .default = VALUE))
```

Calculate pop (= population size) and avg_VALUE (= average VALUE2) by CMA. Set na.rm = TRUE in weighted.mean(). Assign temp as its name.

Drop observation where CMA is "Other census metropolitan areas, census agglomerations and other geographies".

Sort the remaining observations in the descending order of avg_VALUE.
]

---

.tiny[
```{r include=params$answers}
pumf <- read_rds("../Data/2016_census_pumf_ind.rds")
pumf <- pumf %>% mutate(VALUE2 = case_when( VALUE > 80000000 ~ NA, .default = VALUE))

temp <- pumf %>% group_by(CMA) %>% 
  summarise(
    avg_VALUE = weighted.mean(VALUE2, w=WEIGHT, na.rm = TRUE), 
    pop=sum(WEIGHT)
    )

temp <- temp %>% 
  filter(CMA != "Other census metropolitan areas, census agglomerations and other geographies")

temp <- temp %>% arrange( desc(avg_VALUE) )
```
]

---

```{r}
temp
```

---

Create a scatterplot showing the population size on x-axis and the average VALUE2 on y-axis.

---

```{r echo=params$answers}
ggplot() + geom_point(aes(pop, avg_VALUE), temp)
```

---

Install `ggrepel` package, and add the following geom: 

.tiny[
```{r eval=FALSE}
+ ggrepel::geom_text_repel(
  aes(pop, avg_VALUE, label=CMA), 
  temp %>% arrange(desc(pop)) %>% head(8) 
  )
```
]


---

.tiny[
```{r echo=params$answers}
ggplot() + geom_point(aes(pop, avg_VALUE), temp) + 
  ggrepel::geom_text_repel(
    aes(pop, avg_VALUE, label=CMA), 
    temp %>% arrange(desc(pop)) %>% head(8) )
```
]
