---
title: "ggplot2 - 2"
subtitle: "Aesthetic"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = NA, fig.asp = 0.56, out.width = "70%", dpi = 300, fig.align = "center")
options(tibble.print_max = Inf)
library(kableExtra)
```

aesthetics

- color
- size
- shape
- fill
- group
- linetype

attribute

scale functions

---

```{r}
library(lubridate)
library(tidyverse)
```

```{r}
bca <- read_rds("Data/bca252.rds")
```

---

```{r}
by_aa <- bca %>%
  filter(!is.na(aa_label)) %>%
  group_by(aa_label) %>%
  summarise(
    avg_floorarea = mean(floorarea, na.rm = TRUE),
    avg_price = mean(price, na.rm = TRUE),
    n = n()
  )
```

---

#### Practice - Create the following figures.

.small[
```{r echo=FALSE}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color = aa_label, size = n))
```
]

--

.small[ 
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color = aa_label, size = n))
```
]

---

```{r echo=FALSE}
ggplot(by_aa) +
  geom_col(aes(aa_label, n, fill = aa_label))
```

--

.small[ 
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa) +
  geom_col(aes(aa_label, n, fill = aa_label))
```
]

---

.small[
```{r}
by_aa_yr <- bca %>%
  filter(!is.na(aa_label)) %>%
  group_by(aa_label, year = year(sale_date)) %>%
  summarise(avg_price = mean(price, na.rm = TRUE), n = n())
```
]

```{r eval=FALSE}
by_aa_yr
```

.scroll-output[
```{r echo=FALSE}
by_aa_yr %>% kable()
```
]

---

line plot by aa_label

.small[
```{r echo=FALSE}
ggplot(by_aa_yr) +
  geom_line(aes(year, avg_price, group = aa_label))
```
]

--

.small[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa_yr) +
  geom_line(aes(year, avg_price, group = aa_label))
```
]

---

.small[
```{r echo=FALSE}
ggplot(by_aa_yr) +
  geom_line(aes(year, avg_price, color = aa_label, linetype = aa_label))
```
]

--

.small[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa_yr) +
  geom_line(aes(year, avg_price, color = aa_label, linetype = aa_label))
```
]

---

### Attributes

```{r}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color = "blue"))
```

---

.small[
```{r}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color = "Sanghoon"))
```
]

---

```{r}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price), color = "blue")
```

---

Generate this figure.

```{r echo=FALSE}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price), color = "blue", size = 5)
```

--

```{r echo=params$answers, eval=FALSE}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price), color = "blue", size = 5)
```

---

Over-plotting

```{r}
van17 <- bca %>% filter(aa_id == 9, year(sale_date) == 2017)
ggplot(van17) +
  geom_point(aes(floorarea, price))
```

---

Generate this figure using alpha aesthetic.

```{r echo=FALSE}
ggplot(van17) +
  geom_point(aes(floorarea, price), alpha = 0.1)
```

--

```{r echo=params$answers, eval=FALSE}
ggplot(van17) +
  geom_point(aes(floorarea, price), alpha = 0.1)
```

---

### Scale functions

.tiny[
```{r}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color = aa_label, size = n))
```
]

---

We want to

- Remove the size legend.

- Rename the name (aa_label) for the color legend to "Assessment Area".

- Format numbers on axis better. (e.g., $900K instead of 9e+05, 2,200 instead of 2200.)

- Rename axis names.

Note these are all related to aesthetic properties.

--

Use scale functions!

---

.tiny[
```{r}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color = aa_label, size = n)) +
  scale_color_discrete(name = "Assessment Area") +
  scale_size(guide = NULL)
```
]

---

.horizontal-scroll[
.tiny[
```{r}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color = aa_label, size = n)) +
  scale_color_discrete(name = "Assessment Area") +
  scale_size(guide = NULL) +
  scale_x_continuous(name = "Average Floor Area (sq ft)", labels = scales::label_comma()) +
  scale_y_continuous(name = "Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K"))
```
]
]

---

Generate the figure on next slide, by building on the following code: 

.tiny[
```{r}
ggplot(by_aa_yr, aes(year, avg_price, color = aa_label)) +
  geom_line() + geom_point()
```
]

---

```{r echo=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color = aa_label)) +
  geom_line() + geom_point() +
  scale_color_discrete(name = "Assessment Area") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M"))
```
--

.scroll-output[
.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color = aa_label)) +
  geom_line() + geom_point() +
  scale_color_discrete(name = "Assessment Area") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M"))
```
]
]

---

.small[
Generate the figure on next slide, by modifying the code below. (Use log scale for y-axis and use breaks argument in its scale function to control the numbers displayed on y-axis.) 
]

.tiny[
.scroll-output[
```{r }
ggplot(by_aa_yr, aes(year, avg_price, color = aa_label)) +
  geom_line() + geom_point() +
  scale_color_discrete(name = "Assessment Area") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M"))
```
]
]

---

```{r echo=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color = aa_label)) +
  geom_line() + geom_point() +
  scale_color_discrete(name = "Assessment Area") +
  scale_x_continuous(name = "Year") +
  scale_y_log10(
    name = "Average Price",
    labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M"),
    breaks = c(200000, 300000, 500000, 700000, 1000000, 1500000, 2000000)
  )
```

--

.scroll-output[
.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color = aa_label)) +
  geom_line() + geom_point() + scale_color_discrete(name = "Assessment Area") + scale_x_continuous(name = "Year") +
  scale_y_log10( name = "Average Price",labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M"), 
  							 breaks = c(200000, 300000, 500000, 700000, 1000000, 1500000, 2000000))
```
]
]


---

Over-plotting continued

```{r}
ggplot(van17) +
  geom_point(aes(floorarea, price), alpha = 0.1)
```

---

Zoom in

.tiny[
```{r}
ggplot(van17) +
  geom_point(aes(floorarea, price), alpha = 0.1) +
  scale_x_continuous(limits = c(200, 3000)) +
  scale_y_continuous(limits = c(0, 5000000))
```
]

---

Use log scale

```{r}
ggplot(van17) +
  geom_point(aes(floorarea, price), alpha = 0.1) +
  scale_x_log10() + scale_y_log10()
```

---

Generate the figure on next slide, by building on the following code:

.tiny[
```{r}
library(gapminder)
gapminder_2007 <- gapminder %>% filter(year == 2007)
ggplot(gapminder_2007) +
  geom_point(aes(x = gdpPercap, y = lifeExp, color = continent))
```
]

---

```{r echo=FALSE}
ggplot(gapminder_2007) +
  geom_point(aes(x = gdpPercap, y = lifeExp, color = continent)) +
  scale_x_log10(name = "GDP per capita", labels = scales::label_comma()) +
  scale_y_continuous(name = "Life expectancy") +
  scale_color_discrete(name = "Continent")
```

--

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(gapminder_2007) +
  geom_point(aes(x = gdpPercap, y = lifeExp, color = continent)) +
  scale_x_log10(name = "GDP per capita", labels = scales::label_comma()) +
  scale_y_continuous(name = "Life expectancy") +
  scale_color_discrete(name = "Continent")
```
]

