---
title: "ggplot2 - 3"
subtitle: "Geometries"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

Labeling points in a scatter plot

Bar plots

Histogram

---

.small[
```{r results='hide'}
library(tidyverse)
library(gapminder)
gapminder_2007 <- gapminder %>% filter(year == 2007)
gapminder_2007
```
]

.scroll-output[
```{r echo=FALSE}
gapminder_2007 %>% knitr::kable()
```
]

---

#### Labeling points

.tiny[
```{r}
gapminder_2007_big <- gapminder_2007 %>% top_n(20, pop)

ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + geom_point() + 
  geom_text(aes(label=country), gapminder_2007_big)+ scale_x_log10()
```
]

---

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + geom_point() + 
  ggrepel::geom_text_repel(aes(label=country), gapminder_2007_big) + 
  scale_x_log10()
```

---

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + geom_point() + 
  ggrepel::geom_label_repel(aes(label=country), gapminder_2007_big) + 
  scale_x_log10()
```

---

Practice tasks

```{r}
bca <- read_rds("Data/bca252.rds")

by_aa <- bca %>% 
  filter(! is.na(aa_label)) %>% 
  group_by(aa_label) %>% 
  summarise(
    avg_floorarea = mean(floorarea, na.rm = TRUE), 
    avg_price=mean(price, na.rm=TRUE), 
    n=n())

```

---

.small[
Generate the figure on next slide, by building on the following codes:
]

.tiny[
```{r}
ggplot(by_aa) +
  geom_point(aes(avg_floorarea, avg_price, color=aa_label, size=n)) +
  scale_color_discrete(name="Assessment Area") +   scale_size(guide=NULL) +
  scale_x_continuous(name="Average Floor Area (sq ft)", labels = scales::label_comma()) +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K"))
```
]

---

```{r echo=FALSE}
ggplot(by_aa, aes(avg_floorarea, avg_price, color=aa_label)) +
  geom_point(aes(size=n)) +
  ggrepel::geom_text_repel(aes(label=aa_label)) +
  scale_color_discrete(guide=NULL) + 
  scale_size(guide=NULL) +
  scale_x_continuous(name="Average Floor Area (sq ft)", labels = scales::label_comma()) +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K"))
```
--

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa, aes(avg_floorarea, avg_price, color=aa_label)) +
  geom_point(aes(size=n)) + ggrepel::geom_text_repel(aes(label=aa_label)) +
  scale_color_discrete(guide=NULL) + scale_size(guide=NULL) +
  scale_x_continuous(name="Average Floor Area (sq ft)", labels = scales::label_comma()) +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K"))
```
]

---

### Barchart

```{r}
bca_2017 <- read_rds("Data/bca252.rds") %>% 
  filter(lubridate::year(sale_date) == 2017)
```

---

```{r}
aa_n <- bca_2017 %>% group_by(aa_label) %>% summarise(n=n())
```

```{r echo=FALSE}
aa_n %>% kable
```

---

```{r}
ggplot(aa_n) + geom_col(aes(aa_label, n))
```

---

```{r}
ggplot(bca_2017) + geom_bar(aes(aa_label))
```

---

.scroll-output[
.tiny[
```{r echo=FALSE}
bca_2017 %>% head(20) %>% kable
```
]]

---

.tiny[
```{r}
bca_2017_aa_act <- bca_2017 %>% group_by(aa_label, act_use_cat) %>% 
	summarise(n=n(), avg_price = mean(price, na.rm=TRUE))
```
]

.scroll-output[
.tiny[
```{r echo=FALSE}
bca_2017_aa_act %>% head(20) %>% kable
```
]]


---

.tiny[
```{r}
ggplot(bca_2017_aa_act) + geom_col(aes(aa_label, n, fill=act_use_cat), position = "stack")
```
]

---

```{r}
ggplot(bca_2017_aa_act) + 
  geom_col(aes(aa_label, n, fill=act_use_cat), position = "fill")
```

---

```{r}
ggplot(bca_2017_aa_act) + 
  geom_col(aes(aa_label, n, fill=act_use_cat), position = "dodge")
```

---

`geom_bar()` automatically calculates counts and use them as y. But `geom_col()` is more flexible. 

```{r}
ggplot(bca_2017_aa_act) + 
  geom_col(aes(aa_label, avg_price, fill=act_use_cat), position = "dodge")
```

---

Default value for `position=`

[? geom_col()](https://ggplot2.tidyverse.org/reference/geom_bar.html)
--

```{r}
ggplot(bca_2017_aa_act) + geom_col(aes(aa_label, n, fill=act_use_cat))
```

---

```{r}
ggplot(bca_2017_aa_act) + geom_col(aes(aa_label, n, fill=act_use_cat)) + coord_flip()
```

---

Practice task - Generate this figure, using `bca_2017_aa_act`.

```{r echo=FALSE}
ggplot(bca_2017_aa_act) + 
  geom_col(aes(act_use_cat, avg_price, fill=aa_label), position = "dodge")
```

--

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(bca_2017_aa_act) + geom_col(aes(act_use_cat, avg_price, fill=aa_label), position = "dodge")
```
]

---

Practice task

```{r}
pumf_prihm <- read_rds("Data/2021_census_pumf_ind.rds") %>% 
  filter(PRIHM == "Person is primary maintainer", 
         AGEGRP != "Not available", 
         Tenur != "Not available")
```

Generate figures on next two slides. Try both `geom_col()` and `geom_bar()`.

---
 
```{r echo=FALSE}
ggplot(pumf_prihm) + geom_bar(aes(AGEGRP, fill=Tenur, weight=WEIGHT), position = "fill") + coord_flip()
```
--

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(pumf_prihm) + 
  geom_bar(aes(AGEGRP, fill=Tenur, weight=WEIGHT), position = "fill") + 
  coord_flip()
```
]

---

```{r echo=FALSE}
ggplot(pumf_prihm) + geom_bar(aes(AGEGRP, fill=Tenur, weight=WEIGHT), position = "fill") + coord_flip() +
  scale_x_discrete(name=NULL) + 
  scale_y_continuous(name = NULL, labels= scales::label_number(scale=100, suffix = '%')) +
  scale_fill_discrete(name= NULL, labels= c("Own", "Rent"))
```
--
.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(pumf_prihm) + geom_bar(aes(AGEGRP, fill=Tenur, weight=WEIGHT), position = "fill") + coord_flip() +
  scale_x_discrete(name=NULL) + 
  scale_y_continuous(name = NULL, labels= scales::label_number(scale=100, suffix = '%')) +
  scale_fill_discrete(name= NULL, labels= c("Own", "Rent"))
```
]

---

### Histogram

```{r}
ggplot(bca_2017) + geom_histogram(aes(price))
```

---

```{r}
ggplot(bca_2017) + geom_histogram(aes(price)) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price), binwidth = 1000) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price), binwidth = 1000) + 
  scale_x_continuous(limits = c(400000,450000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price), binwidth = 100000) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 100000) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

.tiny[
```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 100000, position = "dodge") + 
  scale_x_continuous(limits = c(0,6000000))
```
]

---

.tiny[
```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 1000000, position = "fill") + 
  scale_x_continuous(limits = c(0,10000000))
```
]

---

### Practice task

Build on the last code to create this plot.
```{r echo=FALSE}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 1000000, position = "fill") + 
  scale_x_continuous(name = "Price", limits = c(0,10000000), labels = scales::label_comma(scale=0.000001, prefix = "$", suffix = "M")) +
  scale_y_continuous(name = NULL, labels = scales::label_comma(scale=100, suffix = "%")) +
  scale_fill_discrete(name = NULL, labels = c("Single family", "Condo or \nTownhouse")) 
```

---

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 1000000, position = "fill") + 
  scale_x_continuous(name = "Price", limits = c(0,10000000), labels = scales::label_comma(scale=0.000001, prefix = "$", suffix = "M")) +
  scale_y_continuous(name = NULL, labels = scales::label_comma(scale=100, suffix = "%")) +
  scale_fill_discrete(name = NULL, labels = c("Single family", "Condo or \nTownhouse")) 
```
]

