---
title: "ggplot2 - 4"
subtitle: "Theme, facet_wrap, labs"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

- theme
    - customize the non-data components of your plots: i.e. fonts, background, gridlines, legend style, etc.
    - default themes
    - legend.position
    - text = element_text(size=  )

- facet_wrap

- labs

---

```{r}
library(tidyverse)

bca <- read_rds("Data/bca252.rds")

avg_price <- bca %>% 
  mutate(sale_year = year(sale_date)) %>% 
  group_by(act_use_cat, sale_year) %>% 
  summarise(avg_price = mean(price, na.rm = TRUE))
```

---

.scroll-output[
```{r echo=FALSE}
avg_price %>% head(30) %>% kable()
```
]

---

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +
	scale_color_discrete(name=NULL) +
	scale_x_continuous(name="Year") +
	scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K"))
```
]

---

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +
	scale_color_discrete(name=NULL) +
	scale_x_continuous(name="Year") +
	scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K")) +
	theme_bw()
```
]

---

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +
	scale_color_discrete(name=NULL) +
	scale_x_continuous(name="Year") +
	scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K")) +
	theme_bw() + theme(legend.position = c(0.2, 0.8))
```
]

---

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +
	scale_color_discrete(name=NULL) +
	scale_x_continuous(name="Year") +
	scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K")) +
	theme_bw() + theme(legend.position = c(0.2, 0.8), text=element_text(size=14))
```
]

---

Practice Task

.tiny[
```{r}
by_aa_yr <- bca %>% 
  filter(! is.na(aa_label)) %>% 
  group_by(aa_label, year=year(sale_date)) %>% 
  summarise(avg_price = mean(price, na.rm = TRUE), n=n())
```
]
.scroll-output[
.tiny[
```{r echo=FALSE}
by_aa_yr %>% head(10) %>% kable
```
]]

---

Create the figure on next slide. Try the themes in the cheatsheet.

.tiny[
```{r}
ggplot(by_aa_yr, aes(year, avg_price, color=aa_label)) + geom_line() + geom_point() +
  scale_color_discrete(name="Assessment Area") + 
  scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M"))
```
]

---

```{r echo=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color=aa_label)) + geom_line() + geom_point() +
  scale_color_discrete(name=NULL) + 
  scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M")) +
	theme_minimal() +
	theme(legend.position = c(0.25, 0.7), text=element_text(size=12))
```

--

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color=aa_label)) + geom_line() + geom_point() +
  scale_color_discrete(name=NULL) + 
  scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M")) +
	theme_minimal() +	theme(legend.position = c(0.25, 0.7), text=element_text(size=12))
```
]

---

facet_wrap()

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +
	scale_color_discrete(name=NULL) +
	scale_x_continuous(name="Year") +
	scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K"))
```
]

---

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price)) + 
  geom_line() + geom_point() +
	scale_color_discrete(name=NULL) +
	scale_x_continuous(name="Year") +
	scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K")) +
	facet_wrap(~act_use_cat)
```
]

---

Practice task

.tiny[
```{r}
ggplot(by_aa_yr, aes(year, avg_price, color=aa_label)) + geom_line() + geom_point() +
  scale_color_discrete(name="Assessment Area") + 
  scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M"))
```
]

---

```{r echo=FALSE}
ggplot(by_aa_yr, aes(year, avg_price)) + geom_line() + geom_point() +
  scale_color_discrete(name="Assessment Area") + 
  scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M")) + facet_wrap(~aa_label)
```

--

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa_yr, aes(year, avg_price)) + geom_line() + geom_point() +
  scale_color_discrete(name="Assessment Area") + 
  scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M")) + 
	facet_wrap(~aa_label)
```
]

---
labs()

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +	
	scale_color_discrete(name=NULL) +	scale_x_continuous(name="Year") +
	scale_y_continuous(name="Average Price", labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K")) +
	theme_bw() + theme(legend.position = c(0.2, 0.8))
```
]

---

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +	
	scale_y_continuous(labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K")) + 
	labs(color=NULL, x="Year", y="Average Price") +
	theme_bw() + theme(legend.position = c(0.2, 0.8))
```
]

---

.tiny[
```{r}
ggplot(avg_price, aes(sale_year, avg_price, color=act_use_cat)) + 
  geom_line() + geom_point() +	
	scale_y_continuous(labels = scales::label_comma(scale = 0.001, prefix = "$", suffix = "K")) + 
	labs(color=NULL, x="Year", y="Average Price", title = "Average Transanction Price History by Building Type", caption = "Data source: BC Assessment Authority") +
	theme_bw() + theme(legend.position = c(0.2, 0.8))
```
]

---

Practice task

.tiny[
```{r}
ggplot(by_aa_yr, aes(year, avg_price, color=aa_label)) + geom_line() + geom_point() +
  scale_color_discrete(name=NULL) + scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M")) +
	theme_minimal() + theme(legend.position = c(0.25, 0.7), text=element_text(size=12))
```
]

---

```{r echo=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color=aa_label)) + geom_line() + geom_point() +
  scale_color_discrete(name=NULL) + scale_x_continuous(name="Year") +
  scale_y_continuous(name="Average Price", labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M")) +
	labs(title= "Average Transaction Price History By Assessment Area", caption = "Data Source: BC Assessment Authority") +
	theme_minimal() + theme(legend.position = c(0.25, 0.7), text=element_text(size=12)) 
```

--

.tiny[
```{r echo=params$answers, eval=FALSE}
ggplot(by_aa_yr, aes(year, avg_price, color=aa_label)) + geom_line() + geom_point() +
  scale_y_continuous(labels = scales::label_comma(accuracy = 0.1, scale = 0.000001, prefix = "$", suffix = "M")) +
	labs(color=NULL, x="Year", y="Average Price", title= "Average Transaction Price History By Assessment Area", caption = "Data Source: BC Assessment Authority") +
	theme_minimal() + theme(legend.position = c(0.2, 0.7), text=element_text(size=12)) 
```
]
