---
title: "Factor"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

```{r}
library(tidyverse)
```

```{r}
bca2020 <- read_rds("Data/bca252.rds") %>%
	filter(year(sale_date)==2020) 
```

---

```{r}
n_wday <- bca2020 %>% 
	group_by(wday = wday(sale_date, label=TRUE)) %>% 
	summarise(n=n())
```

```{r}
n_wday
```

---

```{r}
ggplot(n_wday) + geom_col( aes(wday, n))
```

--

How does `ggplot()` know the order among the weekdays on x-axis?

---

Factor

```{r}
n_wday$wday %>% class
```

```{r}
n_wday$wday
```

---

```{r eval=FALSE}
n_wday %>% mutate(wday2 = as.numeric(wday))
```

```{r echo=FALSE}
n_wday %>% mutate(wday2 = as.numeric(wday)) %>% kable
```

For a factor, `ggplot` uses the order in the underlying levels as coordinates. 

---

```{r}
n_wday2 <- n_wday %>% mutate(wday = as.character(wday))
```

```{r}
n_wday2
```

---

```{r}
ggplot(n_wday2) + geom_col(aes(wday, n))
```

For a character vector, it uses the alphabetical order for the position.

---

You can create a factor using factor().

.small[
```{r}
tb <- tibble(
	municipality = c("Vancouver", "Richmond", "Surrey", "Port Moody"), 
	value=c(5,1,8,3))
```
]
```{r echo=FALSE}
tb %>% kable
```

---

```{r}
ggplot(tb) + geom_col(aes(x=value, y=municipality))
```

---

.small[
```{r}
tb <- tb %>% mutate(
	municipality2 = factor(municipality, levels=c("Richmond", "Port Moody", "Vancouver", "Surrey")))
```
]
```{r echo=FALSE}
tb %>% kable
```

```{r}
tb$municipality2
```

---

```{r}
ggplot(tb) + geom_col(aes(x=value, y=municipality2))
```

---

```{r}
ggplot(tb) + geom_col(aes(x=value, y=fct_reorder(municipality,value)))
```


---

.small[
```{r}
ggplot(tb) + 
	geom_col(aes(x=value, y=fct_reorder(municipality,value) %>% fct_rev() ))
```
]

---

### Pracice: Create the figure on the following slide.

.small[
```{r}
avgp_aa <- bca2020 %>% group_by(aa_label) %>% 
	summarise(avgp = mean(price))
```
]
```{r echo=FALSE}
avgp_aa %>% kable
```

---

```{r echo=FALSE}
ggplot(avgp_aa) + geom_col(aes(avgp, fct_reorder(aa_label, avgp))) + 
	scale_x_continuous(name="Average Price", labels = scales::label_comma(scale=0.000001, prefix="$", suffix = "M")) + scale_y_discrete(name=NULL)
```

--

.tiny[
```{r eval=FALSE}
ggplot(avgp_aa) + geom_col(aes(avgp, fct_reorder(aa_label, avgp))) + 
	scale_x_continuous(name="Average Price", labels = scales::label_comma(scale=0.000001, prefix="$", suffix = "M")) + scale_y_discrete(name=NULL)
```
]

---

#### Don't use it if the x values have a natural order by themselves.

```{r}
ggplot(n_wday2) + geom_col(aes(fct_reorder(wday,n), n)) 
```

---

### Ordered factor

.small[
```{r}
tb <- tb %>% mutate(municipality3 = factor(municipality, levels=c("Surrey", "Vancouver", "Richmond", "Port Moody"), ordered = TRUE))
```

```{r}
tb$municipality3 %>% class
```

```{r}
tb$municipality3
```
]

---

You can use inequality conditions with an ordered factor.

```{r eval=F}
tb %>% mutate(temp = municipality3 >= "Richmond")
```
```{r echo=F}
tb %>% mutate(temp = municipality3 >= "Richmond") %>% kable
```

---

Suppose you want to combine Vancouver and Richmond to have the same name "Vancouver-Richmond".

```{r eval=FALSE}
tb %>% mutate( municipality4 = case_when(
	municipality3 == "Surrey" ~ "Surrey",
	municipality3 <= "Richmond" ~ "Vancouver-Richmond",
	municipality3 == "Port Moody" ~ "Port Moody",
	.default = NA
))
```

---

```{r echo=FALSE}
tb %>% mutate( municipality4 = case_when(
	municipality3 == "Surrey" ~ "Surrey",
	municipality3 <= "Richmond" ~ "Vancouver-Richmond",
	municipality3 == "Port Moody" ~ "Port Moody",
	.default = NA
)) %>% kable
```

---

Practice

```{r}
pumf <- read_rds("data/2021_census_pumf_ind.rds")
```
```{r}
pumf$HDGREE %>% class
```

---

.scroll-output[
.tiny[
```{r}
pumf$HDGREE	%>% levels()
```
]]

---

.small[
Generate the figure on the following slide.

1. Convert HDGREE to an ordered factor
    ```{r}
    pumf <- pumf %>% mutate(HDGREE = as.ordered(HDGREE))
    ```

2. Create HDGREE2 that simplfies HDGREE to have the following levels:  "Less Than High School", "High School Diploma", "Post-Secondary (No Degree)", "Bachelor's Degree", "Graduate Degree". Use case_when()

3. Convert HDGREE2 into a factor so that it will have the natural order.

4. Filter out the observations where HDGREE2 is missing.
    ```{r eval=FALSE}
    filter(! is.na(HDGREE2))
    ```
]

---

```{r echo=FALSE}
pumf %>% mutate(
	HDGREE2 = case_when(
		HDGREE == "No certificate, diploma or degree" ~ "Less Than High School",
		HDGREE == "High (secondary) school diploma or equivalency certificate" ~ "High School Diploma",
		HDGREE <= "University certificate or diploma below bachelor level" ~ "Post-Secondary (No Degree)",
		HDGREE == "Bachelor's degree" ~ "Bachelor's Degree",
		HDGREE <= "Earned doctorate" ~ "Graduate Degree",
		.default = NA
	)
) %>% mutate(HDGREE2 = factor(HDGREE2, levels=c("Less Than High School", "High School Diploma", "Post-Secondary (No Degree)", "Bachelor's Degree", "Graduate Degree" ))) %>% 
	filter(! is.na(HDGREE2)) %>% 
	ggplot() + geom_bar(aes(y=HDGREE2, weight=WEIGHT))
```

---

.small[
```{r}
pumf <- pumf %>% mutate(
	HDGREE2 = case_when(
		HDGREE == "No certificate, diploma or degree" ~ "Less Than High School",
		HDGREE == "High (secondary) school diploma or equivalency certificate" ~ "High School Diploma",
		HDGREE <= "University certificate or diploma below bachelor level" ~ "Post-Secondary (No Degree)",
		HDGREE == "Bachelor's degree" ~ "Bachelor's Degree",
		HDGREE <= "Earned doctorate" ~ "Graduate Degree",
		.default = NA
	)
)
```
]

---

.small[
```{r}
pumf %>% filter(!is.na(HDGREE2)) %>% ggplot() + 
	geom_bar(aes(y=HDGREE2, weight=WEIGHT))
```
]

---

.tiny[
```{r}
pumf <- pumf %>% mutate(
	HDGREE2 = factor(HDGREE2, levels=c("Less Than High School", "High School Diploma", "Post-Secondary (No Degree)", "Bachelor's Degree", "Graduate Degree" )))
```

```{r}
pumf %>% filter(!is.na(HDGREE2)) %>% ggplot() + 
	geom_bar(aes(y=HDGREE2, weight=WEIGHT))
```
]

---

Create this figure.

```{r echo=FALSE}
ggplot(pumf %>% filter(!is.na(HDGREE2))) + 
	geom_bar(aes(y=HDGREE2, weight=WEIGHT, fill=Gender), position = "fill")
```

--

.tiny[
```{r eval=FALSE}
pumf %>% filter(!is.na(HDGREE2)) %>% ggplot() + 
	geom_bar(aes(y=HDGREE2, weight=WEIGHT, fill=Gender), position = "fill")
```
]

---

.tiny[
```{r echo=FALSE}
ggplot(pumf %>% filter(! is.na(HDGREE2))) + geom_bar(aes(y=HDGREE2, weight=WEIGHT, fill=Gender), position = "fill") +
	geom_vline(xintercept = c(0.45,0.5,0.55), linetype="dashed") +
	coord_cartesian(xlim = c(0.4, 0.6)) +
	scale_y_discrete(name=NULL) +
	scale_x_continuous(name= NULL, labels = scales::label_percent()) +
	scale_fill_discrete(name= NULL)
```
]

--

.tiny[
```{r eval=FALSE}
pumf %>% filter(! is.na(HDGREE2)) %>% 
	ggplot() + geom_bar(aes(y=HDGREE2, weight=WEIGHT, fill=Gender), position = "fill") +
	geom_vline(xintercept = c(0.45,0.5,0.55), linetype="dashed") +
	coord_cartesian(xlim = c(0.4, 0.6)) +
	scale_y_discrete(name=NULL) +
	scale_x_continuous(name= NULL, labels = scales::label_percent()) +
	scale_fill_discrete(name= NULL)
```
]

---

```{r echo=FALSE}
pumf %>% filter(
	! is.na(HDGREE2), 
	MarStH == "Never married (not living common law)",
	CMA %>% str_detect("Vancouver|Toronto")) %>% 
ggplot() + geom_bar(aes(y=HDGREE2, weight=WEIGHT, fill=Gender), position = "fill") +
	geom_vline(xintercept = c(0.45,0.5,0.55), linetype="dashed") +
	coord_cartesian(xlim = c(0.4, 0.6)) +
	scale_y_discrete(name=NULL) +
	scale_x_continuous(name= NULL, labels = scales::label_percent()) +
	scale_fill_discrete(name= NULL) +
	facet_wrap(~CMA, nrow = 3) +
	labs(title = "Gender Distributions by Education Level", 
			 subtitle = "- Singles (Never Married)", 
			 caption = "Source: Census 2021 Public Use Microdata File.")
```

---

.tiny[
```{r eval=FALSE}
pumf %>% filter(
	! is.na(HDGREE2), 
	MarStH == "Never married (not living common law)",
	CMA %>% str_detect("Vancouver|Toronto")) %>% 
ggplot() + geom_bar(aes(y=HDGREE2, weight=WEIGHT, fill=Gender), position = "fill") +
	geom_vline(xintercept = c(0.45,0.5,0.55), linetype="dashed") +
	coord_cartesian(xlim = c(0.4, 0.6)) +
	scale_y_discrete(name=NULL) +
	scale_x_continuous(name= NULL, labels = scales::label_percent()) +
	scale_fill_discrete(name= NULL) +
	facet_wrap(~CMA, nrow = 3) +
	labs(title = "Gender Distributions by Education Level", 
			 subtitle = "- Singles (Never Married)", 
			 caption = "Source: Census 2021 Public Use Microdata File.")
```
]