---
title: "StatCan Data Tables"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

```{r}
library(tidyverse)
```

---

#### Statistics Canada Data Table (Formely, CANSIM)

Go to [Statistics Canada website](https://www12.statcan.gc.ca).

Click Data.

Browse Subject category.

- Which subjects are related to real estate?
    - Browse available tables in such subjects.

---

Browse Geography category.

- Which geographic unit corresponds to municipalities (e.g., city of Vancouver, Burnaby, ...)?

--

Browse Frequency category.

- What frequency has the largest number of available tables except Occasional?

--

Browse Survey or Statistical Programs

- Which programs are related to real estate?

---

#### Vacancy rates

Search for 'vacancy rates' in keyword box.

Click Housing in Subject, Census metropolitan area in Geography.

What is table number for 'Canada Mortgage and Housing Corporation, vacancy rates, apartment structures of six units and over, privately initiated in census metropolitan areas'?

--

34-10-0127-01

---

```{r}
library(cansim)
```

```{r}
vacancy <- get_cansim("34-10-0127-01")
```

---

.scroll-output-80[
.tiny[
```{r echo=FALSE}
vacancy %>% head(20) %>% kable()
```
]]

---

.scroll-output[
.tiny[
```{r}
vacancy$GEO %>% unique()
```
]]

---

.tiny[
```{r out.width='90%'}
vacancy %>% filter( GEO %>% str_detect("Vancouver|Toronto") ) %>% 
	ggplot(aes(Date, VALUE, color=GEO)) + geom_line() + geom_smooth(se=FALSE) +
	labs(title="Vacancy Rates", x=NULL, y=NULL, color=NULL) +
	scale_y_continuous(labels = scales::label_percent(scale = 1 )) +
	theme_bw() + theme(legend.position = "bottom")
```
]

---


#### Practice - Create this graph. (Replace ##)

```{r echo=FALSE, out.width='90%'}
mortgage_rate <- get_cansim("34-10-0145-01")

ggplot(mortgage_rate, aes(Date, VALUE)) + geom_line() +
	labs(title="Conventional mortgage lending rate, 5-year term", caption="Source: Canada Mortgage and Housing Corporation\nStatistics Canada. Table 34-10-0145-##", 
			 x=NULL, y=NULL) +
	scale_y_continuous(labels=scales::label_percent(scale=1))
```

---

.tiny[
```{r eval=FALSE}
mortgage_rate <- get_cansim("34-10-0145-01")

ggplot(mortgage_rate, aes(Date, VALUE)) + geom_line() +
	labs(
		title="Conventional mortgage lending rate, 5-year term", 
		caption="Source: Canada Mortgage and Housing Corporation\nStatistics Canada. Table 34-10-0145-01", 
		 x=NULL, y=NULL) +
	scale_y_continuous(labels=scales::label_percent(scale=1))
```
]

---

### Practice - Create this graph.

```{r echo=FALSE, out.width='90%'}
population <- get_cansim("17-10-0148-01")

population %>% filter(GEO %>% str_detect("Vancouver"), Gender=="Total - gender", `Age group`=="All ages") %>% 
	ggplot(aes(Date, VALUE)) + geom_line() + geom_point() +
	labs(title = "Vancouver Metro Population Trend", x=NULL, y=NULL, caption = "Annual, July 1 estimates, y-axis in log scale.\nSource: Statistics Canada Table: 17-10-0148-01") + 
	scale_y_log10(labels=scales::label_number(scale = 0.000001, suffix="M")) + theme_bw()
```

---

.tiny[
```{r eval=FALSE}
population <- get_cansim("17-10-0148-01")

population %>% filter(GEO %>% str_detect("Vancouver"), Gender=="Total - gender", `Age group`=="All ages") %>% 
	ggplot(aes(Date, VALUE)) + geom_line() + geom_point() +
	labs(title = "Vancouver Metro Population Trend", x=NULL, y=NULL, caption = "Annual, July 1 estimates, y-axis in log scale.\nSource: Statistics Canada Table: 17-10-0148-01") + 
	scale_y_log10(labels=scales::label_number(scale = 0.000001, suffix="M")) + theme_bw()
```
]

---

.small[
```{r eval=FALSE}
population %>% view()
```
]
.scroll-output-80[
.tiny[
```{r echo=FALSE}
population %>% head(200) %>% kable()
```
]]

---

.scroll-output[
.tiny[
```{r}
population$GEO %>% unique()
```
]]

---

.scroll-output[
.tiny[
```{r}
population$Gender %>% unique()
```
]]

---
.scroll-output[
.tiny[
```{r}
population$`Age group` %>% unique()
```
]]

---

.tiny[
```{r eval=FALSE}
population <- get_cansim("17-10-0148-01")

population %>% filter(
	GEO %>% str_detect("Vancouver"), 
	Gender=="Total - gender",
	`Age group`=="All ages") %>% 
	ggplot(aes(Date, VALUE)) + geom_line() + geom_point() +
	labs(
		title = "Vancouver Metro Population Trend", 
		x=NULL, y=NULL, 
		caption = "Annual, July 1 estimates, y-axis in log scale.\nSource: Statistics Canada Table: 17-10-0148-01") + 
	scale_y_log10(labels=scales::label_number(scale = 0.000001, suffix="M")) + theme_bw()
```
]
