---
title: "Census profile"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
options(tibble.print_max = Inf, message = FALSE, warn = FALSE)
library(kableExtra)
```

```{r}
library(tidyverse)
library(cancensus)
```

---

Census public use microdata file (PUMF) publishes data at individual level. 

Census Profile publishes data at geographic area level (e.g., whole Canada, province, CMA, CSD, tract).

- The smallest geographic unit for PUMF is CMA.

- Use PUMF if the unit of your analysis is individuals. Use Profile if it is geographic areas.

To see the type of available information at a geographic level, go to the '[Census Profile, 2021 Census](https://www12.statcan.gc.ca/census-recensement/2021/dp-pd/prof/index.cfm?Lang=E)' page at the Statistics Canada website (Google it). Click 'Canada' in the Browse box. 

---

Most of the data are available at each of the Census geographic levels.

.small[
<table>
<thead>
<tr class="header">
<th>Code</th>
<th>Description</th>
<th align="center">Count in Census 2016</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td>C</td>
<td>Canada (total)</td>
<td align="center">1</td>
</tr>
<tr class="even">
<td>PR</td>
<td>Provinces/Territories</td>
<td align="center">13</td>
</tr>
<tr class="odd">
<td>CMA</td>
<td>Census Metropolitan Area</td>
<td align="center">35</td>
</tr>
<tr class="even">
<td>CA</td>
<td>Census Agglomeration</td>
<td align="center">14</td>
</tr>
<tr class="odd">
<td>CD</td>
<td>Census Division</td>
<td align="center">287</td>
</tr>
<tr class="even">
<td>CSD</td>
<td>Census Subdivision</td>
<td align="center">713</td>
</tr>
<tr class="odd">
<td>CT</td>
<td>Census Tracts</td>
<td align="center">5621</td>
</tr>
<tr class="even">
<td>DA</td>
<td>Dissemination Area</td>
<td align="center">56589</td>
</tr>
<tr class="odd">
<td>EA</td>
<td>Enumeration Area (1996 only)</td>
<td align="center">-</td>
</tr>
<tr class="even">
<td>DB</td>
<td>Dissemination Block (2001-2021)</td>
<td align="center">489676</td>
</tr>
<!-- <tr class="odd"> -->
<!-- <td>Regions</td> -->
<!-- <td>Named Census Region</td> -->
<!-- <td align="center"></td> -->
<!-- </tr> -->
</tbody>
</table>
]

---

```{r echo=FALSE, fig.dim=c(9,9)}
csd <- get_census(dataset = "CA21", regions = list(CMA = "59933"), level = "CSD", use_cache = TRUE, geo_format = "sf")
ggplot(csd) +
  geom_sf() +
  geom_sf_text(aes(label = name %>% str_extract(".*(?=\\(..\\))")), data = csd %>% filter(name %>% str_detect("(CY)|(DM)|(IM)")), size = 5) +
  theme_void() +
  labs(title = "CSD (Census Subdivision) boundaries") +
  theme(plot.title = element_text(hjust = 0.5))
```

---

```{r echo=FALSE, fig.dim=c(9,9)}
ct <- get_census(dataset = "CA21", regions = list(CMA = "59933"), level = "CT", use_cache = TRUE, geo_format = "sf")
ggplot(ct) +
  geom_sf() +
  ggtitle("Census tract boundaries") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

---

We will learn how to import Census profile data into R, using `cancensus` package. For example, 

.scroll-output[
.tiny[
```{r echo=FALSE}
get_census("CA21", list(C = "01"), level = "CMA", vectors = "v_CA21_4311") %>%
  arrange(desc(Population)) %>%
  select(matches("GeoUID|Name|^v_")) %>%
  filter(!is.na(`v_CA21_4311: Median value of dwellings ($) (60)`)) %>%
  kable()
```
]]

---

.scroll-output[
.tiny[
```{r echo=FALSE}
get_census("CA21", list(CMA = "59933"), level = "CT", vectors = "v_CA21_4311") %>%
  select(matches("GeoUID|^v_")) %>%
  filter(!is.na(`v_CA21_4311: Median value of dwellings ($) (60)`)) %>%
  head(100) %>%
  kable()
```
]]

---

.small[It is difficult to interpret data at small geographic units. We will learn how to show them on a map.]
```{r echo=FALSE, fig.dim=c(9,9), fig.align='center'}
get_census("CA21", list(CMA = "59933"), level = "CT", vectors = "v_CA21_4311", geo_format = "sf") %>%
  select(matches("GeoUID|^v_")) %>%
  filter(!is.na(`v_CA21_4311: Median value of dwellings ($) (60)`)) %>%
  ggplot() +
  geom_sf(aes(fill = `v_CA21_4311: Median value of dwellings ($) (60)`)) +
  scale_fill_viridis_c(name = "Median dwelling value\n (Census Tract)", labels = scales::label_dollar(scale = 0.000001, suffix = "M")) +
  theme_void()
```

---

```{r eval=FALSE}
install.packages("cancensus")
install.packages("sf")
```

```{r eval=FALSE}
library(cancensus)
library(tidyverse)
```

---

Click [here](https://censusmapper.ca/users/sign_up) to sign up for Censusmapper.ca to obtain your own API key. You can use your Google, Facebook or Twitter account.

Run the following code <i>in Console just once</i>, to store your API key in your computer and automatically use it in API calls. 
```{r eval=FALSE}
set_cancensus_api_key("your_api_key", install = TRUE)
```

Run the following code <i>in Console just once</i>, to set the cache folder.
```{r eval=FALSE}
set_cancensus_cache_path("~/cancensus_cache", install = TRUE)
```

---

Suppose we want to calculate Chinese shares by CMA.

```{r}
chinese_cma <- get_census(
  dataset = "CA21", regions = list(C = "01"),
  vectors = c("v_CA21_4872", "v_CA21_4881"), level = "CMA"
)
```

This code downloads total population and the number of Chinese by CMA.

The following columns are included by default: Area (sq km), Population, Dwellings, Households.

---

.scroll-output[
.tiny[
```{r echo=FALSE}
chinese_cma %>% kable()
```
]]

---

.tiny[
```{r eval=FALSE}
chinese_cma %>%
  transmute(
    `Region Name`,
    Chinese_share = `v_CA21_4881: Chinese` / `v_CA21_4872: Total - Visible minority for the population in private households`
  ) %>%
  arrange(desc(Chinese_share))
```
]
.scroll-output[
.tiny[
```{r echo=FALSE}
chinese_cma %>%
  transmute(`Region Name`, Chinese_share = `v_CA21_4881: Chinese` / `v_CA21_4872: Total - Visible minority for the population in private households`) %>%
  arrange(desc(Chinese_share)) %>%
  kable()
```
]]

---

```{r eval=FALSE}
chinese_cma <- get_census(
  dataset = "CA21",
  regions = list(C = "01"),
  vectors = c("v_CA21_4872", "v_CA21_4881"),
  level = "CMA"
)
```

---

#### dataset="CA21"

```{r eval=FALSE}
list_census_datasets() %>% view()
```

.scroll-output[
.tiny[
```{r echo=FALSE}
list_census_datasets() %>% kable()
```
]]

---

#### regions= list(C="01")

Used for geographic coverage

```{r eval=FALSE}
list_census_regions("CA21") %>% view()
```

.scroll-output[
.tiny[
```{r echo=FALSE}
list_census_regions("CA21") %>%
  head(100) %>%
  kable()
```
]]

---

.small[
Whole Canada
```{r eval=FALSE}
regions <- list(C = "01")
```

Vancouver CMA
```{r eval=FALSE}
regions <- list(CMA = "59933")
```

Vancouver or Toronto CMAs
```{r eval=FALSE}
regions <- list(CMA = c("59933", "35535"))
```

Vancouver or Toronto CMAs or Alberta
```{r eval=FALSE}
regions <- list(CMA = c("59933", "35535"), PR = "48")
```
]

---

#### level="CMA"

.small[
<table>
<thead>
<tr class="header">
<th>Code</th>
<th>Description</th>
<th align="center">Count in Census 2016</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td>C</td>
<td>Canada (total)</td>
<td align="center">1</td>
</tr>
<tr class="even">
<td>PR</td>
<td>Provinces/Territories</td>
<td align="center">13</td>
</tr>
<tr class="odd">
<td>CMA</td>
<td>Census Metropolitan Area</td>
<td align="center">35</td>
</tr>
<tr class="even">
<td>CA</td>
<td>Census Agglomeration</td>
<td align="center">14</td>
</tr>
<tr class="odd">
<td>CD</td>
<td>Census Division</td>
<td align="center">287</td>
</tr>
<tr class="even">
<td>CSD</td>
<td>Census Subdivision</td>
<td align="center">713</td>
</tr>
<tr class="odd">
<td>CT</td>
<td>Census Tracts</td>
<td align="center">5621</td>
</tr>
<tr class="even">
<td>DA</td>
<td>Dissemination Area</td>
<td align="center">56589</td>
</tr>
<tr class="odd">
<td>EA</td>
<td>Enumeration Area (1996 only)</td>
<td align="center">-</td>
</tr>
<tr class="even">
<td>DB</td>
<td>Dissemination Block (2001-2016)</td>
<td align="center">489676</td>
</tr>
<!-- <tr class="odd"> -->
<!-- <td>Regions</td> -->
<!-- <td>Named Census Region</td> -->
<!-- <td align="center"></td> -->
<!-- </tr> -->
</tbody>
</table>
]

---

#### vectors= c("v_CA21_4872","v_CA21_4881")

At [the Canada page of Census Profile, 2021 Census website](https://www12.statcan.gc.ca/census-recensement/2021/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=01&Geo2=&Code2=&SearchText=Canada&SearchType=Begins&SearchPR=01&B1=All&TABID=1&type=0), find the characteric label you are looking for.

Note the hierarchy of the characteristics. Chinese < Total visible minority population < Total - Visible minority for the population in private households - 25% sample data

```{r eval=FALSE}
explore_census_vectors("CA21")
```

Type "Chinese" in the Search box.

---

```{r eval=FALSE}
chinese_cma <- get_census(
  dataset = "CA21",
  regions = list(C = "01"),
  vectors = c("v_CA21_4872", "v_CA21_4881"),
  level = "CMA"
)
```


---

### Practice

Calculate Chinese share by CSD in Vancouver CMA. Sort them in the descending order in the share.

---

.small[
The following code downloads total population and the number of Chinese by CSD within Vancouver CMA.

```{r include=params$answers}
chinese_csd_van <- get_census(
  dataset = "CA21", regions = list(CMA = "59933"),
  vectors = c("v_CA21_4872", "v_CA21_4881"), level = "CSD"
)
```
]

---

.scroll-output[
.tiny[
```{r echo=FALSE}
chinese_csd_van %>% kable()
```
]]

---

.tiny[
```{r eval=FALSE}
chinese_csd_van %>%
  transmute(`Region Name`,
    Chinese_share = `v_CA21_4881: Chinese` / `v_CA21_4872: Total - Visible minority for the population in private households`
  ) %>%
  arrange(desc(Chinese_share))
```
]
.scroll-output[
.tiny[
```{r echo=FALSE}
chinese_csd_van %>%
  transmute(`Region Name`,
    Chinese_share = `v_CA21_4881: Chinese` / `v_CA21_4872: Total - Visible minority for the population in private households`
  ) %>%
  arrange(desc(Chinese_share)) %>%
  kable()
```
]]

---

### Practice

Calculate Renter household share by CSD in Vancouver CMA. Sort them by the renter share in descending order. 

```{r include=params$answers, eval=FALSE}
explore_census_vectors("CA21")
```

---

.tiny[
```{r include=params$answers}
renter_csd_van <- get_census("CA21", list(CMA = "59933"),
  level = "CSD",
  vectors = c("v_CA21_4239", "v_CA21_4237")
)
```
]

.scroll-output[
.tiny[
```{r include=params$answers, echo=FALSE}
renter_csd_van %>% kable()
```
]]

---

.tiny[
```{r include=params$answers, eval=FALSE}
renter_csd_van %>%
  transmute(
    `Region Name`,
    renter_share = `v_CA21_4239: Renter` / `v_CA21_4237: Total - Private households by tenure`
  ) %>%
  arrange(desc(renter_share))
```
]

.tiny[
```{r include=params$answers, echo=FALSE}
renter_csd_van %>%
  transmute(
    `Region Name`,
    renter_share = `v_CA21_4239: Renter` / `v_CA21_4237: Total - Private households by tenure`
  ) %>%
  arrange(desc(renter_share)) %>%
  kable()
```
]


---

### Map

```{r}
chinese_csd_van_sf <- get_census(
  dataset = "CA21",
  regions = list(CMA = "59933"),
  vectors = c("v_CA21_4872", "v_CA21_4881"),
  level = "CSD",
  geo_format = "sf"
)
```

geo_format = 'sf' create an sf object, which extends data.frame-like objects with the geometry column.


```{r}
class(chinese_csd_van_sf)
```

---

.scroll-output[
.tiny[
```{r echo=FALSE}
chinese_csd_van_sf %>% kable()
```
]]

---

.tiny[
```{r}
chinese_share_csd_van_sf <- chinese_csd_van_sf %>%
  transmute(
    `Region Name`,
    Chinese_share = `v_CA21_4881: Chinese` / `v_CA21_4872: Total - Visible minority for the population in private households`
  ) %>%
  arrange(desc(Chinese_share))
```
]
---

You can use `geom_sf()` to show an sf object as a map.

.tiny[
```{r}
ggplot(chinese_share_csd_van_sf) +
  geom_sf(aes(fill = Chinese_share)) +
  theme_void()
```
]

---

#### Practice

Show the renter share at CSD level, on a map.

.tiny[
```{r include=params$answers}
renter_csd_van_sf <- get_census("CA21", list(CMA = "59933"),
  level = "CSD",
  vectors = c("v_CA21_4239", "v_CA21_4237"), geo_format = "sf"
)
```
]

---

```{r include=params$answers}
renter_share_csd_van_sf <- renter_csd_van_sf %>%
  transmute(
    `Region Name`,
    renter_share = `v_CA21_4239: Renter` / `v_CA21_4237: Total - Private households by tenure`
  ) %>%
  arrange(desc(renter_share))
```

---

.tiny[
```{r include=params$answers}
ggplot(renter_share_csd_van_sf) +
  geom_sf(aes(fill = renter_share)) +
  theme_void()
```
]
---

### Practice at home

Show the Chinese share at Census tract level

--

.tiny[
```{r include=params$answers, eval=FALSE}
chinese_ct_van_sf <- get_census(
  dataset = "CA21",
  regions = list(CMA = "59933"),
  vectors = c("v_CA21_4872", "v_CA21_4881"),
  level = "CT",
  geo_format = "sf"
)

chinese_share_ct_van_sf <- chinese_ct_van_sf %>%
  transmute(Chinese_share = `v_CA21_4881: Chinese` / `v_CA21_4872: Total - Visible minority for the population in private households`) %>%
  arrange(desc(Chinese_share))

ggplot(chinese_share_ct_van_sf) +
  geom_sf(aes(fill = Chinese_share)) +
  theme_void()
```
]

---

```{r include=params$answers, echo=FALSE, fig.dim=c(9,9)}
chinese_ct_van_sf <- get_census(
  dataset = "CA21",
  regions = list(CMA = "59933"),
  vectors = c("v_CA21_4872", "v_CA21_4881"),
  level = "CT",
  geo_format = "sf"
)

chinese_share_ct_van_sf <- chinese_ct_van_sf %>%
  transmute(Chinese_share = `v_CA21_4881: Chinese` / `v_CA21_4872: Total - Visible minority for the population in private households`) %>%
  arrange(desc(Chinese_share))

ggplot(chinese_share_ct_van_sf) +
  geom_sf(aes(fill = Chinese_share)) +
  theme_void()
```


---

Show the renter share at Census tract level

--

.tiny[
```{r include=params$answers, eval=FALSE, fig.dim=c(9,9)}
renter_ct_van_sf <- get_census("CA21", list(CMA = "59933"),
  level = "CT",
  vectors = c("v_CA21_4239", "v_CA21_4237"), geo_format = "sf"
)

renter_share_ct_van_sf <- renter_ct_van_sf %>%
  transmute(
    `Region Name`,
    renter_share = `v_CA21_4239: Renter` / `v_CA21_4237: Total - Private households by tenure`
  ) %>%
  arrange(desc(renter_share))

ggplot(renter_share_ct_van_sf) +
  geom_sf(aes(fill = renter_share)) +
  theme_void()
```
]

---

```{r include=params$answers, echo=FALSE, fig.dim=c(9,9)}
renter_ct_van_sf <- get_census("CA21", list(CMA = "59933"),
  level = "CT",
  vectors = c("v_CA21_4239", "v_CA21_4237"), geo_format = "sf"
)

renter_share_ct_van_sf <- renter_ct_van_sf %>%
  transmute(
    `Region Name`,
    renter_share = `v_CA21_4239: Renter` / `v_CA21_4237: Total - Private households by tenure`
  ) %>%
  arrange(desc(renter_share))

ggplot(renter_share_ct_van_sf) +
  geom_sf(aes(fill = renter_share)) +
  theme_void()
```
