---
title: "Simple features 1"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

```{r}
library(tidyverse)
library(cancensus)
library(sf)
```

---

Obtain the number of commuters and that of public transit users among the commuters for each CT in Vancouver CMA along with geographic boundaries, as a simple feature collection.

```{r}
tb_public_transit <- get_census("CA21", 
                                c("v_CA21_7632","v_CA21_7644"), 
                                regions = list(CMA = "59933"), 
                                level = "CT", 
                                geo_format = "sf")
```

`geo_format = 'sf'` creates a simple feature collection, which extends data.frame-like objects with the geometry column.

---

A simple feature collection has the geometry column, which contains geographic information for each observation. It can be points, lines, or polygons.

The geometry column is sticky.

```{r eval=FALSE}
tb_public_transit %>% select(GeoUID, starts_with("v_CA21")) %>% view
```

.scroll-output[
.tiny[
```{r echo=FALSE}
tb_public_transit %>% select(GeoUID, starts_with("v_CA21")) %>% head(20) %>% kable()
```
]]

---

Use `geom_sf()` to show a simple feature collection.

```{r out.width= '100%'}
ggplot(tb_public_transit) + geom_sf() + theme_void()
```

---

.tiny[
```{r out.width='85%'}
ggplot(tb_public_transit) + 
  geom_sf(aes(fill = `v_CA21_7644: Public transit` / `v_CA21_7632: Total - Main mode of commuting for the employed labour force aged 15 years and over with a usual place of work or no fixed workplace address`)) + 
  scale_fill_viridis_c(name="Public transit\nuser share") + theme_void()
```
]

---

#### Practice

```{r}
csd_bndry <- get_census( "CA21",  
                         regions = list(CMA = "59933"), 
                         level = "CSD", 
                         geo_format = "sf")
```

1. Confirm that csd_bndry is a sf collection.
1. Generate the map on the following slide.

---

```{r echo=FALSE, out.width='100%'}
ggplot(tb_public_transit) + geom_sf(aes(fill = `v_CA21_7644: Public transit` / `v_CA21_7632: Total - Main mode of commuting for the employed labour force aged 15 years and over with a usual place of work or no fixed workplace address`)) + geom_sf(data=csd_bndry, color="lightgrey", alpha=0) + 
  scale_fill_viridis_c(name="Public transit\nuser share", label=scales::number_format(accuracy = 1, scale=100, suffix = "%")) + theme_void()
```


---

.tiny[
```{r eval=FALSE, echo=params$answers}
ggplot(tb_public_transit) + 
	geom_sf(aes(fill = `v_CA21_7644: Public transit` / `v_CA21_7632: Total - Main mode of commuting for the employed labour force aged 15 years and over with a usual place of work or no fixed workplace address`)) +
	geom_sf(data=csd_bndry, color="grey", alpha=0) + 
	scale_fill_viridis_c(
		name="Public transit\nuser share", 
		label=scales::number_format(accuracy = 1, scale=100, suffix = "%")) + 
	theme_void()
```
]

---

#### Local Government Open Data Portals

Go to [City of Vancouver Open Data Portal](https://opendata.vancouver.ca/pages/home/) by Googling it.

Click [Explore the catalogue](https://opendata.vancouver.ca/explore/) and browse available data.
- Many data sets are available as maps.

Other municipalities and provincial governments also run their data portals.

- [DataBC](https://www.data.gov.bc.ca/)
- [Burnaby](https://data.burnaby.ca/)
- [Surrey](https://data.surrey.ca/)

---

#### VancouvR

```{r}
library(VancouvR)
```

For smaller datasets, no API key is needed, but for larger ones, you must [register at the City of Vancouver Open Data Portal](https://opendata.vancouver.ca/signup/) to get an API key.

```{r eval=FALSE}
options(VancouverOpenDataApiKey="your api key")
```

---

Search for [Rapid transit stations](https://opendata.vancouver.ca/explore/dataset/rapid-transit-stations/information/) at the City of Vancouver open data website.

Note the Dataset Identifier.

```{r}
stations <- get_cov_data("rapid-transit-stations", format = "geojson")
```

`format = "geojson"` gives a simple feature collection.

Do the same for [Rapid transit lines](https://opendata.vancouver.ca/explore/dataset/rapid-transit-lines/information/)

```{r}
lines <- get_cov_data("rapid-transit-lines", format = "geojson")
```

Confirm that both `stations` and `lines` are sf collections with point and line types.

---

.scroll-output[
.tiny[
```{r echo=FALSE}
stations %>% knitr::kable()
```
]]

---

.scroll-output[
.tiny[
```{r echo=FALSE}
lines %>% knitr::kable()
```
]]

---

#### Practice - Create this map, using `geom_sf()`, `geom_sf_text()`:

```{r out.width='100%', echo=FALSE}
ggplot() + geom_sf(data = lines) + 
  geom_sf(data = stations, size=3, color="red") + 
  geom_sf_text(aes(label=station), data = stations, size = 2) + 
  theme_void()
```

---

```{r out.width='100%', eval=FALSE}
ggplot() + geom_sf(data = lines) + 
  geom_sf(data = stations, size=3, color="red") + 
  geom_sf_text(aes(label=station), data = stations, size = 2) + 
  theme_void()
```

---

#### Adding a ggmap background 

.tiny[
```{r}
library(ggmap)
van <- get_map("Queen Elizabeth park, Vancouver", zoom = 12, 
               source = "stadia", maptype = "stamen_toner_lite")
ggmap(van)
```
]

---

- To use a ggmap background with a sf related geometric function, you have to add `inherit.aes = FALSE` in each geometric function.
- To zoom in, use `coord_sf()`.

```{r eval=FALSE}
ggmap(van) + geom_sf(data = lines, inherit.aes = FALSE) + 
  geom_sf(data = stations, size=3, color="red", inherit.aes = FALSE) + 
  geom_sf_text(aes(label=station), data = stations, size = 2, inherit.aes = FALSE) + 
  coord_sf(ylim = c(49.2,49.3)) + 
	theme_void()
```

---

```{r echo=FALSE, out.width="100%"}
ggmap(van) + geom_sf(data = lines, inherit.aes = FALSE) + 
  geom_sf(data = stations, size=3, color="red", inherit.aes = FALSE) + 
  geom_sf_text(aes(label=station), data = stations, size = 2, inherit.aes = FALSE) + 
  coord_sf(ylim = c(49.2,49.3)) +
	theme_void()
```

---

#### Practice - Create a zoning map for city of Vancouver

```{r echo=FALSE}
zoning <- get_cov_data("zoning-districts-and-labels", format = "geojson")
```

```{r echo=FALSE, out.width="250%"}
ggmap(van) + geom_sf(aes(fill=zoning_classification), zoning, inherit.aes = FALSE, alpha=0.2) + 
	labs(fill=NULL) + coord_sf(xlim = c(-123.3, -123.02), ylim = c(49.2,49.3)) + theme_void()
```

---

```{r eval=FALSE}
zoning <- get_cov_data("zoning-districts-and-labels", format = "geojson")

ggmap(van) + 
	geom_sf(
		aes(fill=zoning_classification), zoning, 
					inherit.aes = FALSE, alpha=0.2) + 
	labs(fill=NULL) + 
	coord_sf(xlim = c(-123.3, -123.02), ylim = c(49.2,49.3)) + 
	theme_void()
```

---

```{r eval=FALSE}
library(mapview)
mapview(zoning, zcol="zoning_classification")
```

Do not use the interactive mode map to generate a static map!

---

```{r echo=FALSE}
library(mapview)
mapview(zoning, zcol="zoning_classification")
```

---

#### (Optional) Additional resource

Another great R package for creating a map is `tmap` package.

See this online book to learn more about how to make a map using R.
https://bookdown.org/nicohahn/making_maps_with_r5/docs/introduction.html

---

### Importing a Shapefile or GeoJSON as a sf collection

An API package like `VancouvR` is not always available. We need to know how to read standard spatial files into R.

--

Go to [Burnaby's Open Data Portal](https://data.burnaby.ca/). 

Find Park Inventory. 

Download data as a GeoJSON file and move it into data folder.

Read it into R using `st_read()`.
```{r results='hide'}
burnaby_parks <- st_read("Data/Park_Inventory.geojson")
```

---

```{r}
mapview(burnaby_parks, zcol="NAME", legend=FALSE)
```

---

Download the same data as a Shapefile and extract it into data folder.

Read it into R using `st_read()`.
```{r results='hide'}
burnaby_parks_2 <- st_read("Data/Park_Inventory/Park_Inventory.shp")
```

A Shapefile consists of a bundle of files with the same name but with different extensions. Even though you specify the one with .shp in `st_read()`, you need all the associated files.

You should get the same result.

---

```{r}
mapview(burnaby_parks_2, zcol="NAME", legend=FALSE)
```

