---
title: "Simple features 2 - Geoprocessing"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = NA, fig.asp = 0.56, out.width = "70%", dpi = 300, fig.align = "center")
knitr::opts_knit$set(root.dir = "..")
options(tibble.print_max = Inf)
library(kableExtra)
```


Find zoning code for each housing transaction
- Spatial join
- Converting a tibble with coordinates to sf
- CRS

Find housing transactions within 100 metres from any skytrain station
- Buffer


---

```{r}
library(tidyverse)
library(sf)
library(VancouvR)
library(mapview)
```

```{r}
zoning <- get_cov_data("zoning-districts-and-labels") %>%
  select(zoning_classification)
```

---

```{r echo=FALSE}
zoning %>% knitr::kable()
```

---

```{r}
mapview(zoning, zcol = "zoning_classification")
```

---

### Converting a tibble with coordinates to sf

So far we have downloaded sf collections. Another way to obtain a sf collection is to convert a tibble with x and y coordinates. For example, the BC assessment data has `lon` and `lat`.

.small[
```{r}
bca <- read_rds("Data/bca252.rds")

van_2023 <- bca %>%
  filter(
    year(sale_date) == 2023,
    juris_id == 200,
    !is.na(lon),
    !is.na(lat)
  ) %>%
  select(folioid, lon, lat, sale_date, price)
```
]

---

.scroll-output[
.tiny[
```{r echo=FALSE}
van_2023 %>%
  head(100) %>%
  kable()
```
]]

---

#### Coordinate Reference System (CRS)

The system used to express a location on the Earth as coordinates on a map. For example, the [Mercator projection](https://en.wikipedia.org/wiki/Mercator_projection) uses longitude and latitude as X and Y coordinates on a map.

Every CRS has some distortion. For example, the Mercator projection preserves an angle but distorts area. For small areas (e.g., city) the distortion is minimal.

Numerous CRSs have been developed with different properties. We have to choose the best CRS for our purpose.

The EPSG code is a unique identifier assigned to a specific Coordinate Reference System (CRS) by the EPSG (European Petroleum Survey Group) Geodetic Parameter Dataset. 

---

We will use these CRS in this course:

- WGS 84 - [EPSG: 4326](https://epsg.io/4326)
    - Can be used anywhere in the world.
    - Uses longitude and latitude as coordinates.
- UTM zone 10N - [EPSG:32610](https://epsg.io/32610)
    - Area of use: Between 126°W and 120°W, northern hemisphere between equator and 84°N, onshore and offshore. Canada - **British Columbia (BC)**; Northwest Territories (NWT); Nunavut; Yukon. United States (USA) - Alaska (AK).
    - unit: **metre**. Much more convenient than degrees used for long and lat (e.g., distance calculation). 

---

```{r}
van_2023_sf <- st_as_sf(van_2023, coords = c("lon", "lat"), crs = 4326)
```

Confirm `van_2023_sf` is a simple feature collection.
```{r}
van_2023_sf %>% head()
```

---

.scroll-output[
.tiny[
```{r echo=FALSE}
van_2023_sf %>%
  head(100) %>%
  kable()
```
]]

---

#### Why Convert it to a sf collection?

You can process them as data. For example, we can fetch zoning information from the zoning map.

---

```{r}
ggplot() +
  geom_sf(aes(fill = zoning_classification), zoning) +
  geom_sf(data = van_2023_sf) +
  theme_void()
```

---

### Spatial Join

Geoprocessing (e.g., spatial join) requires all sf collections to have the same CRS.
```{r}
st_crs(van_2023_sf) == st_crs(zoning)
```
```{r}
van_2023_sf_zoning <- van_2023_sf %>% st_join(zoning)
```

---

.scroll-output[
.tiny[
```{r echo=FALSE}
van_2023_sf_zoning %>%
  head(100) %>%
  kable()
```
]]

---

#### Practice - Import the VPD data and convert it into an sf collection

Go to [VPD Open data website](https://geodash.vpd.ca/opendata/) and download data for 2023 - All neighbourhoods.

Unzip it and move the csv file into `data` folder.

```{r}
crimes <- read_csv("Data/crimedata_csv_AllNeighbourhoods_2026/crimedata_csv_AllNeighbourhoods_2026.csv")
```

---

Note that X and Y are not long and lat. The pdf manual says it is UTM Zone 10.

.scroll-output-80[
.tiny[
```{r echo=FALSE}
crimes %>%
  head(100) %>%
  kable()
```
]]

---

Remove observations where their precise locations are suppressed: X = 0 or Y = 0.

```{r}
crimes <- crimes %>%
  filter(X != 0, Y != 0)
```

---

Convert crimes into a sf collection and name it `crimes_sf`. 

--

```{r out.width="110%"}
crimes_sf <- crimes %>%
  st_as_sf(coords = c("X", "Y"), crs = 32610)
```

---

```{r}
ggplot(crimes_sf) +
  geom_sf(alpha = 0.1) +
  theme_void()
```

---

### Buffer analysis

Identify transactions within 300 metres from each skytrain station. (We will continue to use `van_2023_sf`.)

```{r}
stations <- get_cov_data("rapid-transit-stations") %>%
  select(station)
```

---

```{r}
ggplot() +
  geom_sf(data = van_2023_sf, alpha = 0.2, size = 0.5) +
  geom_sf(data = stations, color = "brown", size = 2) +
  theme_void()
```

---

Use `st_buffer()`. It pads up a geometric object with a given distance. Google it for illustrations.

We have to transform the CRS of `stations` to UTM zone 10N (EPSG 32610), because it uses metres for distance.

```{r}
stations <- stations %>% st_transform(32610)
stations_300 <- stations %>% st_buffer(300)
```

---

```{r}
ggplot() +
  geom_sf(data = van_2023_sf, alpha = 0.2, size = 0.5) +
  geom_sf(data = stations_300, color = "brown", size = 2, alpha = 0.3) +
  theme_void()
```
---

### Spatial join

Spatial join requires all sf collections to have the same CRS.

```{r}
van_2023_sf_stations_300 <- van_2023_sf %>%
  st_transform(st_crs(stations_300)) %>%
  st_join(stations_300)
```

---

```{r}
van_2023_sf_stations_300 %>% head()
```
---

```{r}
van_2023_sf_stations_300 %>%
  filter(!is.na(station)) %>%
  head()
```

---

```{r}
ggplot(van_2023_sf_stations_300) +
  geom_sf(aes(color = station)) +
  theme_void()
```
---

Calculate the number of transactions within 300 meters from each station

You can convert a simple feature collection to a regular tibble with `st_drop_geometry()`. It drops the geometry column.

```{r}
van_2023_sf_stations_300 %>%
  st_drop_geometry() %>%
  count(station, sort = TRUE) %>%
  head()
```

---

#### Practice - Count # of crimes that happened within 500 metres for each station

Use `crimes_sf` and `stations`.

---

It is usually a good idea to reproject the layers to have the same CRS at the beginning. Since we have to take a 500 metre buffer, we must use UTM Zone 10N (EPSG:32610). But `stations` and `crimes_sf` are already in UTM Zone 10N, so we do not need to transform them.

If they were not, we could use
```{r}
stations <- stations %>% st_transform(32610)
crimes_sf <- crimes_sf %>% st_transform(32610)
```

---

```{r}
stations_500 <- stations %>% st_buffer(500)
crimes_stations_500 <- crimes_sf %>% st_join(stations_500)
crimes_stations_500 %>%
  st_drop_geometry() %>%
  count(station, sort = TRUE)
```

