---
title: "Pivoting"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  answers: TRUE
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

```{r}
library(tidyverse)
```

---

#### pivot_wider()

.tiny[
```{r}
library(cancensus)
renter_csd_van <- get_census("CA16", list(CMA="59933"), level="CSD", 
                             vectors = c("v_CA16_4838", "v_CA16_4836")) 

renter_l <- renter_csd_van %>% select(`Region Name`, `v_CA16_4838: Renter`, `v_CA16_4836: Total - Private households by tenure - 25% sample data`) %>% pivot_longer(-`Region Name`, names_to = "vector")
```
]

---

.scroll-output[
.tiny[
```{r echo=FALSE}
renter_l %>% kable()
```
]]

---

Calculate renter share for each CSD

```{r echo=FALSE}
renter_l %>% pivot_wider(`Region Name`, names_from = "vector") %>% 
  transmute(`Region Name`, renter_share = `v_CA16_4838: Renter`/`v_CA16_4836: Total - Private households by tenure - 25% sample data`) %>% arrange(-renter_share) %>% kable()
```

---

.pull-left[
.tiny[
```{r}
long <- tibble( 
  geo_code=c(100, 100, 101, 101),
  geo_name=c("a", "a", "b", "b"), 
  var_name=c("x", "y", "x", "y"), 
  val=c(5,6,7,8))
```

```{r echo=FALSE}
long %>% kable()
```
]]

.pull-right[
.tiny[
```{r}
wide <- long %>% 
  pivot_wider(
    names_from = var_name, 
    values_from = val
    )
```

```{r echo=FALSE}
wide %>% kable()
```
]]

---

#### Application

Calculate renter share for each CSD

```{r echo=FALSE}
renter_l %>% pivot_wider(`Region Name`, names_from = "vector", values_from = "value") %>% 
  transmute(`Region Name`, renter_share = `v_CA16_4838: Renter`/`v_CA16_4836: Total - Private households by tenure - 25% sample data`) %>% arrange(-renter_share) %>% kable()
```

---

.tiny[
```{r eval=params$answers, echo=params$answers}
renter_w <- renter_l %>% pivot_wider(
  names_from = "vector", 
  values_from = "value")
```
]

.scroll-output[
.tiny[
```{r echo=FALSE, eval=params$answers}
renter_w %>% kable()
```
]]

---

.tiny[
```{r eval=FALSE, echo=params$answers}
renter_w %>% transmute(
  `Region Name`, 
  renter_share = `v_CA16_4838: Renter`/`v_CA16_4836: Total - Private households by tenure - 25% sample data`) %>% 
  arrange( desc(renter_share))
```
]

.scroll-output[
.tiny[
```{r echo=FALSE, eval=params$answers}
renter_w %>% transmute(`Region Name`, renter_share = `v_CA16_4838: Renter`/`v_CA16_4836: Total - Private households by tenure - 25% sample data`) %>% arrange( desc(renter_share)) %>% kable()
```
]]

---

#### pivot_longer()

.tiny[
```{r }
pindex <- cansim::get_cansim("18-10-0205-01")

pindex_w <- pindex %>% 
  filter(`New housing price indexes` == "Total (house and land)", 
         GEO %>% str_detect("Vancouver|Toronto|Montréal|Victoria|Edmonton|Calgary"),
         lubridate::ym(REF_DATE) >= lubridate::ymd("2016-12-01")) %>% 
  select(REF_DATE, GEO, VALUE) %>% 
  pivot_wider(names_from = REF_DATE, values_from = VALUE)
```
]

---

.scroll-output-80[
.tiny[
```{r echo=FALSE}
pindex_w %>% kable()
```
]]

---

Generate this plot.

```{r echo=FALSE}
pindex_w %>% pivot_longer( cols = `2016-12`:`2024-01`, names_to = "month") %>% 
  mutate(month=lubridate::ym(month)) %>% 
  ggplot(aes(month, value, color=GEO)) + geom_line()  + labs(x=NULL, y=NULL, color=NULL) +
  theme_bw()
```

---

### Wide to long

.pull-left[
```{r eval=FALSE}
wide
```

```{r echo=FALSE}
wide %>% kable()
```
]

.pull-right[
```{r}
long2 <- wide %>% pivot_longer(
  cols = x:y, 
  names_to = "var", 
  values_to= "val")
```
```{r echo=FALSE}
long2 %>% kable()
```

]

---

##### Application: generate the following figure

```{r echo=FALSE}
pindex_w %>% pivot_longer( cols = `2016-12`:`2024-01`, names_to = "month") %>% 
  mutate(month=lubridate::ym(month)) %>% 
  ggplot(aes(month, value, color=GEO)) + geom_line()  + labs(x=NULL, y=NULL, color=NULL) +
  theme_bw()
```

---

.tiny[
```{r eval=params$answers, echo=params$answers}
pindex_l <- pindex_w %>% pivot_longer( 
  cols = `2016-12`:`2024-01`, 
  names_to = "month")
```
]

.scroll-output[
.tiny[
```{r echo=FALSE, eval=params$answers}
pindex_l %>% kable()
```
]]


---

.tiny[
```{r eval=FALSE, echo=params$answers}
pindex_l %>% 
  mutate(month=lubridate::ym(month)) %>% 
  ggplot(aes(month, value, color=GEO)) + geom_line()  + 
  labs(x=NULL, y=NULL, color=NULL) + theme_bw()
```
]

.scroll-output[
.tiny[
```{r echo=FALSE, eval=params$answers}
pindex_l %>% 
  mutate(month=lubridate::ym(month)) %>% 
  ggplot(aes(month, value, color=GEO)) + geom_line()  + labs(x=NULL, y=NULL, color=NULL) +
  theme_bw()
```
]]

